---
title: 'Data Quality: Lecture 10'
author: "Philipp Drebes"
date: "`r format(Sys.Date(), format='%d.%m.%Y')`"
output:
  pdf_document: 
    keep_tex: yes
    highlight: pygments
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("readxl")) install.packages("readxl", dependencies=TRUE)
if (!require("DescTools")) install.packages("DescTools", dependencies=TRUE)
```

# Task 1

A data set is given from Health Survey for England
(ukdataservice.ac.uk). The variables are described in a codebook.

Examine these variables in terms of normal distribution and outliers

-   Age\
    There are different defined variables available -- use the most
    appropriate, e.g. in terms of scale level

-   Income\
    There are different defined variables available -- use the most
    appropriate, e.g. in terms of scale level

-   Body height

-   Body weight

-   BMI

```{r}
library(readxl)
library(DescTools)

data_hse <- read_excel("hse2016_eul_small.xlsx")
head(data_hse, 10)

qqnorm(data_hse$Age35g)
qqline(data_hse$Age35g)
hist(data_hse$Age35g)

qqnorm(data_hse$srcin01d)
qqline(data_hse$srcin01d)
hist(data_hse$srcin01d)

qqnorm(data_hse$Height)
qqline(data_hse$Height)
hist(data_hse$Height)

qqnorm(data_hse$Weight)
qqline(data_hse$Weight)
hist(data_hse$Weight)

qqnorm(data_hse$BMI)
qqline(data_hse$BMI)
hist(data_hse$BMI)
```

# Task 2

```{r}
numbers <- c(1000, 10000, 100000, 1000000)

for (i in numbers) {
  x <- rnorm(n = i)
  
  start_time <- Sys.time()
  qqnorm(x)
  qqline(x)
  end_time <- Sys.time()
  print(end_time - start_time)
  
  sprintf('QQ-PLot for %i data points\n', i)
}

```

What is the relationship between sample size and runtime?

The relationship is exponential.
