---
title: 'Data Quality: Lecture 4'
author: "Philipp Drebes"
date: "`r format(Sys.Date(), format='%d.%m.%Y')`"
output:
  pdf_document: 
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("naniar")) install.packages("naniar", dependencies=TRUE)
if (!require("readxl")) install.packages("readxl", dependencies=TRUE)
if (!require("mice")) install.packages("mice", dependencies=TRUE)
```

## Task 01
### Missing pattern / Test whether MCAR is given

A data set *riskfactors* is given.

How does the missing pattern look like?

Are the elements of the data set completely at random (MCAR)?

Apply a MCAR test.

#### Missing pattern


```{r}
library(naniar)
library(mice)

head(riskfactors)
md.pattern(riskfactors)
```

#### MCAR Test

```{r}
mcar_test(riskfactors)
```

##### Interpretation

The null hypothesis in this test is that the data is MCAR.
Given that the p-value of the MCAR test is below 0.05, we can conclude the *riskfactors* data is not missing completely at random.

## Task 02
### Missing analysis

A data set *grades* is given.

There are missings in the data set.

Using R, analyze the missing structure.

Apply single imputation for types "mean" and "sample" and compare them.

With this setting you will get to know the two simplest methods of MICE.


#### Missing pattern

```{r}
library(readxl)

grades <- read_excel('data/grades.xlsx')
head(grades)

md.pattern(grades)
```



#### Imputation


```{r}
grades.imp.mean <- mice(grades, method="mean", m=1, maxit=1)
grades.imp.sample <- mice(grades, method="sample", m=1, maxit=1)

grades.mean <- complete(grades.imp.mean)
grades.sample <- complete(grades.imp.sample)

cat('mean TakeHome before imputation: ', mean(grades$TakeHome, na.rm = TRUE))
cat('mean TakeHome after imputation [mean]: ', mean(grades.mean$TakeHome, na.rm = TRUE))
cat('mean TakeHome after imputation [sample]: ', mean(grades.sample$TakeHome, na.rm = TRUE))

cat('mean Final before imputation:', mean(grades$Final, na.rm = TRUE))
cat('mean Final after imputation [mean]:', mean(grades.mean$Final, na.rm = TRUE))
cat('mean Final after imputation [sample]:', mean(grades.sample$Final, na.rm = TRUE))
```


```{r}
head(grades.mean, n = 10)
head(grades.sample, n = 10)
```



## Task 03
### MCAR, MAR or MNAR?
Develop your own examples of the emergence of the MCAR, MAR, and MNAR.

#### MCAR (Missing Completely At Random)
For example, when data are missing for respondents because of a server error on the system that was running the online survey tool.

#### MAR (Missing At Random)
For example, only younger people have missing values for IQ. In that case the probability of missing data on IQ is related to age.

#### MNAR (Missing Not At Random)
For example, when data are missing on IQ and only the people with low IQ values have missing observations for this variable.

