---
title: 'Data Quality: Lecture 4'
author: "Philipp Drebes"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("naniar")) install.packages("naniar", dependencies=TRUE)
if (!require("readxl")) install.packages("readxl", dependencies=TRUE)
if (!require("mice")) install.packages("mice", dependencies=TRUE)
```

## Task 01
### Missing pattern / Test whether MCAR is given

A data set *riskfactors* is given.

How does the missing pattern look like?

Are the elements of the data set completely at random (MCAR)?

Apply a MCAR test.

#### Missing pattern
```{r}
library(naniar)
library(mice)

head(riskfactors)
md.pattern(riskfactors)
```
#### MCAR Test
```{r}
mcar_test(riskfactors)
```

## Task 02
### Missing analysis

A data set *grades* is given.

There are missings in the data set.

Using R, analyze the missing structure.

Apply single imputation for types "mean" and "sample" and compare them.

With this setting you will get to know the two simplest methods of MICE.

#### Missing pattern
```{r}
library(readxl)

grades <- read_excel('data/grades.xlsx')
head(grades)

md.pattern(grades)

mean(grades$TakeHome, na.rm = TRUE)
mean(grades$Final, na.rm = TRUE)


```

#### Imputation
```{r}

imp_grades_mean <- mice(grades, method="mean", m=1, maxit=1)
imp_grades_mean <- mice(grades, method="sample", m=1, maxit=1)

grades_mean <- complete(imp_grades_mean)

head(grades,10)
head(grades_mean,10)

mean(grades$TakeHome, na.rm = TRUE)
mean(grades_mean$TakeHome, na.rm = TRUE)

mean(grades$Final, na.rm = TRUE)
mean(grades_mean$Final, na.rm = TRUE)
```

## Task 03
### MCAR, MAR or MNAR?
Develop your own examples of the emergence of the MCAR, MAR, and MNAR.

#### MCAR

#### MAR

#### MNAR

